### 2023-08-21 proofs/tex/alg-pos-subtyping-lemmas.tex
WIP: leftovers
WIP: description, notation

\begin{lemma}[Soundness of the Positive Subtyping] \label{lemma:pos-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$, $[[Γ ; Θ ⊢ uP]]$, and 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$,
    then $[[Θ ⊢ SC]]$ and
    for any normalized $[[uσ]]$ such that $[[Θ ⊢ uσ : SC]]$,
    $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
\end{lemma}
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$. 
    Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
    $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

@@

    It is left to show that $[[Γ ⊢ [uσ]â⁺ ≥ iP']]$ for any normalized $[[uσ]]$ 
    s.t. $[[Θ ⊢ uσ : (â⁺ :≥ iQ')]]$.
    The latter means that $[[ Θ(â⁺) ⊢ [uσ]â⁺ ≥ iQ' ]]$, i.e. $[[Δ ⊢ [uσ]â⁺ ≥ iQ']]$. 
    By weakening the context to $[[Γ]]$ and combining this judgment
    transitively with $[[Γ ⊢ iQ' ≥ iP]]$, we have $[[Γ ⊢ [uσ]â⁺ ≥ iP]]$,
    as required.

@@

    \item \label{case:pos-subt-soundness:shift} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ SC]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$,
    and it returns the resulting unification constraint $[[lift UC]] = [[SC]]$ as the result.
    By the soundness of unification (\cref{lemma:unification-soundness}),
    $[[Θ ⊢ SC]]$ and for any normalized $[[uσ]]$, $[[Θ ⊢ uσ : SC]]$
    implies $[[ [uσ]nf(uN) = nf(iM) ]]$, 
    then we rewrite the left-hand side by \cref{lemma:norm-subst-distr}:
    $[[ nf([uσ]uN) = nf(iM) ]]$ and apply \cref{lemma:subt-equiv-algorithmization}:
    $[[Γ ⊢ [uσ]uN ≈ iM]]$, then by \ruleref{\ottdruleDOneShiftULabel},
    $[[Γ ⊢ ↓[uσ]uN ≥ ↓iM]]$.

@@

        Let us take an arbitrary normalized $[[uσ]]$ s.t. $[[Θ ⊢ uσ : SC' \ {α̂⁻*}]]$.
        We wish to show $[[Γ ⊢ [uσ]uP ≥ iQ]]$, i.e. $[[Γ ⊢ ∃nas.[uσ]uP' ≥ ∃nbs.iQ']]$.
        To do that, we apply \ruleref{\ottdruleDOneExistsLabel}, and what is left to show is
        $[[Γ, nbs ⊢ [iNs/nas][uσ]uP' ≥ iQ']]$ for some $[[iNs]]$.
        If we construct a normalized $[[uσ']]$ such that $[[Θ, â⁻*[G, nbs] ⊢ uσ' : SC']]$
        and for some $[[iNs]]$, $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$,
        we can apply the induction hypothesis to 
        $[[Γ, nbs; Θ, â⁻*[G, nbs] ⊨ [â⁻*/nas] uP ≥ iQ ⫤ SC']]$ and infer 
        the required subtyping.

@@

        It is easy to see that $[[uσ']]$ is normalized.
        Let us show that $[[Θ, â⁻*[G, nbs] ⊢ uσ' : SC']]$.
        Let us take an arbitrary entry $[[scE]]$ from $[[SC']]$ restricting a variable $[[β̂±]]$.
        Suppose $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. Then
        $[[ (Θ, â⁻*[G, nbs])(β̂±) ⊢ [uσ']β̂± : scE ]]$ is
        rewritten as $[[ Θ(β̂±) ⊢ [uσ]β̂± : scE ]]$, which holds since $[[Θ ⊢ uσ : SC']]$.
        Suppose $[[β̂±]] = [[αî⁻]] \in [[α̂⁻*]]$. Then
        $[[scE]] = [[(αî⁻ :≈ iN)]]$ for some $[[iN]]$, 
        $[[ [uσ']αî⁻ ]] = [[ nf(iN) ]]$ by the definition,
        and $[[ Γ, nbs ⊢ nf(iN) : (αî⁻ :≈ iN) ]]$ by \ruleref{\ottdruleSATSCENEqLabel},
        since $[[Γ ⊢ nf(iN) ≈ iN]]$ by \cref{lemma:subt-equiv-algorithmization}.

        Finally, let us show that $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$.
        For $[[iNi]]$, we take the \emph{normalized} type restricting $[[αî⁻]]$ in $[[SC']]$.
        Let us take an arbitrary variable from $[[uP]]$.
        \begin{enumerate}
            \item If this variable is a unification variable $[[β̂±]]$, then
                $[[ [iNs/nas][uσ] β̂± ]] = [[ [uσ]β̂± ]] $, since $[[Θ ⊢ uσ : SC' \ {α̂⁻*}]]$ and 
                $[[ dom(Θ) ∩ {nas} = ∅ ]]$.

@@

\begin{lemma}[Completeness of the Positive Subtyping] \label{lemma:pos-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$ and $[[Γ ; Θ ⊢ uP]]$.
    Then for any $[[Θ ⊢ uσ]]$ such that $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$,
    there exists $[[Γ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and moreover, $[[Θ ⊢ uσ : SC]]$.
\end{lemma}
\begin{proof}
    Let us prove this lemma by induction on $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation,
    but first, consider the base case for the substitution $[[ [uσ]uP ]]$:
    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ 
            (for potentially empty $[[nbs]]$)\\
            Then by assumption, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ (where $[[ {nbs} ∩ fv [uσ]α̂⁺ = ∅]]$).
            Let us decompose $[[iQ]]$ as $[[iQ]] = [[∃ncs.iQ0]]$, where $[[iQ0]]$ does
            not start with $[[∃]]$.

            By inversion, $[[Γ ; Θ ⊢ ∃nbs.α̂⁺]]$ implies $[[â⁺[Δ] ∊ Θ]]$ for some 
            $[[{Δ} ⊆ {Γ}]]$.

@@

            Moreover, the completeness of upgrade also says that $[[iQ0']]$ is 
            \emph{the least} supertype of $[[iQ0]]$ among types well-formed in $[[Δ]]$, 
            that is $[[Δ ⊢ [uσ]α̂⁺ ≥ iQ0']]$, which means 
            $[[Θ ⊢ uσ : (α̂⁺ :≥ iQ0')]]$, that is $[[Θ ⊢ uσ : SC]]$.

@@

        The algorithm applies \ruleref{\ottdruleAPVarLabel} and 
        infers $[[SC]] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$.
        Then $[[Θ ⊢ uσ : ·]]$ holds trivially.

@@

        \begin{itemize}
            \item $[[nf(uN)]]$ and $[[nf(iM)]]$ are normalized 
            \item $[[Γ ; Θ ⊢ nf(uN)]]$ because $[[Γ ; Θ ⊢ uN]]$ (\cref{todo})
            \item $[[Γ ⊢ nf(iM)]]$ because $[[Γ ⊢ iM]]$ (\cref{corollary:wf-nf})
            \item $[[ Θ ⊢ nf(uσ) ]]$ because $[[Θ ⊢ uσ ]]$ (\cref{todo})
            \item $ \begin{aligned}[t]
                    [[ Γ ⊢ [uσ]uN ≈ iM ]] &\Rightarrow [[ [uσ]uN ≈ iM ]]
                                          && \text {by \cref{lemma:equiv-completeness}}\\
                                          &\Rightarrow [[ nf([uσ]uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:normalization-completeness}}\\
                                          &\Rightarrow [[ [nf(uσ)]nf(uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:norm-subst-distr}}\\
                    \end{aligned}
                  $
        \end{itemize}
        Then by the completeness of the unification,
        $[[Γ ; Θ ⊨ uN ≈u iM ⫤ UC]]$ exists, and
        $[[Θ ⊢ nf(uσ) : lift UC]]$.  Then by \cref{todo}, $[[Θ ⊢ uσ : lift UC]]$.

@@

      To show that the recursive call terminates and that 
      $[[Θ ⊢ uσ : SC1 \ {â⁻*}]]$,
      it suffices to build $[[Θ, â⁻*[G, nbs] ⊢ us0]]$---an extension of $[[uσ]]$ with
      $[[â⁻*]]$ such that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
      Then by the induction hypothesis, $[[Θ, â⁻*[G, nbs] ⊢ us0 : SC1]]$,
      and hence, $[[Θ ⊢ uσ : SC1 \ {â⁻*}]]$, as required.

>>>>

\begin{lemma}[Soundness of the Positive Subtyping] \label{lemma:pos-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$, $[[Γ ; dom(Θ) ⊢  uP]]$, and 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$,
    then $[[Θ ⊢ SC]]$ and
    for any normalized $[[uσ]]$ such that $[[(Θ  ⊢  SC) ⊢ uσ]]$,
    $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
\end{lemma}
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$. 
    Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
    $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

@@

    It is left to show that $[[Γ ⊢ [uσ]â⁺ ≥ iP']]$ for any normalized $[[uσ]]$ 
    s.t. $[[(Θ  ⊢  (â⁺ :≥ iQ')) ⊢ uσ]]$.
    The latter means that $[[ Θ(â⁺) ⊢ [uσ]â⁺ ≥ iQ' ]]$, i.e. $[[Δ ⊢ [uσ]â⁺ ≥ iQ']]$. 
    By weakening the context to $[[Γ]]$ and combining this judgment
    transitively with $[[Γ ⊢ iQ' ≥ iP]]$, we have $[[Γ ⊢ [uσ]â⁺ ≥ iP]]$,
    as required.

@@

    \item \label{case:pos-subt-soundness:shift} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ SC]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$,
    and it returns the resulting unification constraint $[[lift UC]] = [[SC]]$ as the result.
    By the soundness of unification (\cref{lemma:unification-soundness}),
    $[[Θ ⊢ SC]]$ and for any normalized $[[uσ]]$, $[[(Θ  ⊢  SC) ⊢ uσ]]$
    implies $[[ [uσ]nf(uN) = nf(iM) ]]$, 
    then we rewrite the left-hand side by \cref{lemma:norm-subst-distr}:
    $[[ nf([uσ]uN) = nf(iM) ]]$ and apply \cref{lemma:subt-equiv-algorithmization}:
    $[[Γ ⊢ [uσ]uN ≈ iM]]$, then by \ruleref{\ottdruleDOneShiftULabel},
    $[[Γ ⊢ ↓[uσ]uN ≥ ↓iM]]$.

@@

        Let us take an arbitrary normalized $[[uσ]]$ s.t. $[[(Θ  ⊢  SC' \ {α̂⁻*}) ⊢ uσ]]$.
        We wish to show $[[Γ ⊢ [uσ]uP ≥ iQ]]$, i.e. $[[Γ ⊢ ∃nas.[uσ]uP' ≥ ∃nbs.iQ']]$.
        To do that, we apply \ruleref{\ottdruleDOneExistsLabel}, and what is left to show is
        $[[Γ, nbs ⊢ [iNs/nas][uσ]uP' ≥ iQ']]$ for some $[[iNs]]$.
        If we construct a normalized $[[uσ']]$ such that $[[(Θ, â⁻*[G, nbs] ⊢ SC') ⊢ uσ']]$
        and for some $[[iNs]]$, $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$,
        we can apply the induction hypothesis to 
        $[[Γ, nbs; Θ, â⁻*[G, nbs] ⊨ [â⁻*/nas] uP ≥ iQ ⫤ SC']]$ and infer 
        the required subtyping.

@@

        It is easy to see that $[[uσ']]$ is normalized: it inherits this property from 
        $[[uσ]]$.
        Let us show that $[[(Θ, â⁻*[G, nbs]  ⊢ SC') ⊢ uσ' ]] $.
        Let us take an arbitrary entry $[[scE]]$ from $[[SC']]$ restricting a variable $[[β̂±]]$.
        Suppose $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. Then
        $[[ (Θ, â⁻*[G, nbs])(β̂±) ⊢ [uσ']β̂± : scE ]]$ is
        rewritten as $[[ Θ(β̂±) ⊢ [uσ]β̂± : scE ]]$, which holds since $[[(Θ  ⊢  SC') ⊢ uσ]]$.
        Suppose $[[β̂±]] = [[αî⁻]] \in [[α̂⁻*]]$. Then
        $[[scE]] = [[(αî⁻ :≈ iN)]]$ for some $[[iN]]$, 
        $[[ [uσ']αî⁻ ]] = [[ nf(iN) ]]$ by the definition,
        and $[[ Γ, nbs ⊢ nf(iN) : (αî⁻ :≈ iN) ]]$ by \ruleref{\ottdruleSATSCENEqLabel},
        since $[[Γ ⊢ nf(iN) ≈ iN]]$ by \cref{lemma:subt-equiv-algorithmization}.

        Finally, let us show that $[[ [iNs/nas][uσ]uP' ]] = [[ [uσ'][â⁻*/nas]uP' ]]$.
        For $[[iNi]]$, we take the \emph{normalized} type restricting $[[αî⁻]]$ in $[[SC']]$.
        Let us take an arbitrary variable from $[[uP]]$.
        \begin{enumerate}
            \item If this variable is a unification variable $[[β̂±]]$, then
                $[[ [iNs/nas][uσ] β̂± ]] = [[ [uσ]β̂± ]] $, since $[[(Θ  ⊢  SC' \ {α̂⁻*}) ⊢ uσ]]$ and 
                $[[ dom(Θ) ∩ {nas} = ∅ ]]$.

@@

\begin{lemma}[Completeness of the Positive Subtyping] \label{lemma:pos-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$ and $[[Γ ; dom(Θ) ⊢  uP]]$.
    Then for any $[[Θ ⊢ uσ]]$ such that $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$,
    there exists $[[Γ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and moreover, $[[(Θ  ⊢  SC) ⊢ uσ]]$.
\end{lemma}
\begin{proof}
    Let us prove this lemma by induction on $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation,
    but first, consider the base case for the substitution $[[ [uσ]uP ]]$:
    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ 
            (for potentially empty $[[nbs]]$)\\
            Then by assumption, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ (where $[[ {nbs} ∩ fv [uσ]α̂⁺ = ∅]]$).
            Let us decompose $[[iQ]]$ as $[[iQ]] = [[∃ncs.iQ0]]$, where $[[iQ0]]$ does
            not start with $[[∃]]$.

            By inversion, $[[Γ ; dom(Θ) ⊢  ∃nbs.α̂⁺]]$ implies $[[â⁺[Δ] ∊ Θ]]$ for some 
            $[[{Δ} ⊆ {Γ}]]$.

@@

            Moreover, the completeness of upgrade also says that $[[iQ0']]$ is 
            \emph{the least} supertype of $[[iQ0]]$ among types well-formed in $[[Δ]]$, 
            that is $[[Δ ⊢ [uσ]α̂⁺ ≥ iQ0']]$, which means 
            $[[(Θ  ⊢  (α̂⁺ :≥ iQ0')) ⊢ uσ]]$, that is $[[(Θ  ⊢  SC) ⊢ uσ]]$.

@@

        The algorithm applies \ruleref{\ottdruleAPVarLabel} and 
        infers $[[SC]] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$.
        Then $[[(Θ  ⊢  ·) ⊢ uσ]]$ holds trivially.

@@

        \begin{itemize}
            \item $[[nf(uN)]]$ and $[[nf(iM)]]$ are normalized 
            \item $[[Γ ; dom(Θ) ⊢  nf(uN)]]$ because $[[Γ ; dom(Θ) ⊢  uN]]$ (\cref{corollary:wf-nf-algo})
            \item $[[Γ ⊢ nf(iM)]]$ because $[[Γ ⊢ iM]]$ (\cref{corollary:wf-nf})
            \item $[[ Θ ⊢ nf(uσ) ]]$ because $[[Θ ⊢ uσ ]]$ (\cref{corollary:norm-subst-sig-algo})
            \item $ \begin{aligned}[t]
                    [[ Γ ⊢ [uσ]uN ≈ iM ]] &\Rightarrow [[ [uσ]uN ≈ iM ]]
                                          && \text {by \cref{lemma:equiv-completeness}}\\
                                          &\Rightarrow [[ nf([uσ]uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:normalization-completeness}}\\
                                          &\Rightarrow [[ [nf(uσ)]nf(uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:norm-subst-distr}}\\
                    \end{aligned}
                  $
        \end{itemize}
        Then by the completeness of the unification,
        $[[Γ ; Θ ⊨ uN ≈u iM ⫤ UC]]$ exists, and
        $(Θ  ⊢ lift UC) ⊢ nf(uσ)$.  Then by \cref{corollary:sat-equiv}, $[[(Θ  ⊢  lift UC) ⊢ uσ]]$.

@@

      To show that the recursive call terminates and that 
      $[[(Θ  ⊢  SC1 \ {â⁻*}) ⊢ uσ]]$,
      it suffices to build $[[Θ, â⁻*[G, nbs] ⊢ us0]]$---an extension of $[[uσ]]$ with
      $[[â⁻*]]$ such that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
      Then by the induction hypothesis, $[[(Θ, â⁻*[G, nbs] ⊢ SC1) ⊢ us0]]$,
      and hence, $[[(Θ  ⊢  SC1 \ {â⁻*}) ⊢ uσ]]$, as required.

### 2023-08-21 proofs/tex/constraint-merge-lemmas.tex
WIP: description, notation

\begin{lemma} [Soundness of Constraint Merge] \label{lemma:merge-soundness}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$ 
    and $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
    Then 
    \begin{enumerate}
        \item $[[Θ ⊢ SC]]$,
        \item for any substitution $[[Θ ⊢ uσ]]$, $[[Θ ⊢ uσ : SC]]$ implies $[[Θ ⊢ uσ : SC1]]$ and $[[Θ ⊢ uσ : SC2]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    By definition, $[[SC1 & SC2]] = [[SC]]$ consists of three parts:
    entries of $[[SC1]]$ that do not have matching entries of $[[SC]]$,
    entries of $[[SC2]]$ that do not have matching entries of $[[SC1]]$,
    and the merge of matching entries.

@@

    Let us show the second property.
    We take an arbitrary $[[uσ]]$ such that $[[Θ ⊢ uσ]]$ and $[[Θ ⊢ uσ : SC]]$.
    To prove $[[Θ ⊢ uσ : SC1]]$, we need to show that for any $[[scE1]] \in [[SC1]]$, restricting $[[α̂±]]$,
    $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$ holds.

    Let us assume that $[[α̂±]] \notin [[dom(SC2)]]$. It means that $[[SC]] \ni [[scE1]]$, 
    and then since $[[Θ ⊢ uσ : SC]]$, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$.

    Otherwise, $[[SC2]]$ contains an entry $[[scE2]]$ restricting $[[α̂±]]$,
    and $[[SC]] \ni [[scE]]$ where $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$.
    Then since $[[Θ ⊢ uσ : SC]]$, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE]]$,
    and by \cref{lemma:entry-merge-soundness}, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$.

    The proof of $[[Θ ⊢ uσ : SC2]]$ is symmetric.
\end{proof}

@@

\begin{lemma} [Completeness of Constraint Merge] 
    \label{lemma:merge-completeness}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.
    Then for any substitution $[[Θ ⊢ uσ]]$ such that $[[Θ ⊢ uσ : SC1]]$ and $[[Θ ⊢ uσ : SC2]]$, 
    $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
\end{lemma}
\begin{proof}
    By  definition, $[[SC1 & SC2]]$ is a union of
    \begin{enumerate}
        \item entries of $[[SC1]]$, which do not have matching entries in $[[SC2]]$,
        \item entries of $[[SC2]]$, which do not have matching entries in $[[SC1]]$, and 
        \item the merge of matching entries.
    \end{enumerate}

@@

    It is clear that the first two components of this union exist. 
    Let us show that the third component exists.  
    Let us take two entries $[[scE1]] \in [[SC1]]$ and $[[scE2]] \in [[SC2]]$ restricting the same variable $[[α̂±]]$.  $[[Θ ⊢ uσ : SC1]]$ means that $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$ and $[[Θ ⊢ uσ : SC2]]$ means $[[Θ(α̂±) ⊢ [uσ]α̂± : scE2]]$.
    Then by \cref{lemma:entry-merge-completeness}, $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$ is defined and $[[Θ(α̂±) ⊢ [uσ]α̂± : scE]]$.
\end{proof}

\begin{lemma}[Substitution existence]
    \label{lemma:substitution-existence}
    If $[[Θ ⊢ SC]]$ then there exists $[[Θ ⊢ uσ]]$ such that $[[Θ ⊢ uσ : SC]]$.
\end{lemma}
\begin{proof}

>>>>

\begin{lemma} [Soundness of Constraint Merge] \label{lemma:merge-soundness}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$ 
    and $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
    Then 
    \begin{enumerate}
        \item $[[Θ ⊢ SC]]$,
        \item for any substitution $[[Θ ⊢ uσ]]$, $[[(Θ  ⊢  SC) ⊢ uσ]]$ implies $[[(Θ  ⊢  SC1) ⊢ uσ]]$ and $[[(Θ  ⊢  SC2) ⊢ uσ]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    By definition, $[[SC1 & SC2]] = [[SC]]$ consists of three parts:
    entries of $[[SC1]]$ that do not have matching entries of $[[SC]]$,
    entries of $[[SC2]]$ that do not have matching entries of $[[SC1]]$,
    and the merge of matching entries.

@@

    Let us show the second property.
    We take an arbitrary $[[uσ]]$ such that $[[Θ ⊢ uσ]]$ and $[[(Θ  ⊢  SC) ⊢ uσ]]$.
    To prove $[[(Θ  ⊢  SC1) ⊢ uσ]]$, we need to show that for any $[[scE1]] \in [[SC1]]$, restricting $[[α̂±]]$,
    $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$ holds.

    Let us assume that $[[α̂±]] \notin [[dom(SC2)]]$. It means that $[[SC]] \ni [[scE1]]$, 
    and then since $[[(Θ  ⊢  SC) ⊢ uσ]]$, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$.

    Otherwise, $[[SC2]]$ contains an entry $[[scE2]]$ restricting $[[α̂±]]$,
    and $[[SC]] \ni [[scE]]$ where $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$.
    Then since $[[(Θ  ⊢  SC) ⊢ uσ]]$, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE]]$,
    and by \cref{lemma:entry-merge-soundness}, $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$.

    The proof of $[[(Θ  ⊢  SC2) ⊢ uσ]]$ is symmetric.
\end{proof}

@@

\begin{lemma} [Completeness of Constraint Merge] 
    \label{lemma:merge-completeness}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.
    Then for any substitution $[[Θ ⊢ uσ]]$ such that $[[(Θ  ⊢  SC1) ⊢ uσ]]$ and $[[(Θ  ⊢  SC2) ⊢ uσ]]$, 
    $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
\end{lemma}
\begin{proof}
    By  definition, $[[SC1 & SC2]]$ is a union of
    \begin{enumerate}
        \item entries of $[[SC1]]$, which do not have matching entries in $[[SC2]]$,
        \item entries of $[[SC2]]$, which do not have matching entries in $[[SC1]]$, and 
        \item the merge of matching entries.
    \end{enumerate}

@@

    It is clear that the first two components of this union exist. 
    Let us show that the third component exists.  
    Let us take two entries $[[scE1]] \in [[SC1]]$ and $[[scE2]] \in [[SC2]]$ restricting the same variable $[[α̂±]]$.  $[[(Θ  ⊢  SC1) ⊢ uσ]]$ means that $[[Θ(α̂±) ⊢ [uσ]α̂± : scE1]]$ and $[[(Θ  ⊢  SC2) ⊢ uσ]]$ means $[[Θ(α̂±) ⊢ [uσ]α̂± : scE2]]$.
    Then by \cref{lemma:entry-merge-completeness}, $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$ is defined and $[[Θ(α̂±) ⊢ [uσ]α̂± : scE]]$.
\end{proof}

\begin{lemma}[Substitution existence]
    \label{lemma:substitution-existence}
    If $[[Θ ⊢ SC]]$ then there exists $[[Θ ⊢ uσ]]$ such that $[[(Θ  ⊢  SC) ⊢ uσ]]$.
\end{lemma}
\begin{proof}
